---
title: "fishbase_data_in"
output: html_document
date: "2023-04-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(rfishbase)
library(naniar)
```

### Read in data

```{r}
data <- read_csv(here::here("01_data", "02_processed_data", "homerange_pld.csv")) %>%
  mutate(home_range = case_when(
    !is.na(observed_homerange) ~ observed_homerange,
    is.na(observed_homerange) ~ predicted_homerange
  )) %>%
  mutate(pld = case_when(
    !is.na(observed_pld) ~ observed_pld,
    is.na(observed_pld) ~ predicted_pld
  )) %>%
  filter(pld < 400) %>%
  mutate(magnitude_homerange = floor(log10(home_range))) %>%
  mutate(month_pld = ceiling(pld / 30))
```

### Make new clusters

```{r}

clusters <- read_csv(here::here("01_data", "02_processed_data", "clusters.csv")) %>% 
  mutate(class2 = class) %>% 
  mutate(class2 = case_when(class2 == 6 ~ 2, class2 != 6 ~ class2)) # combine classes 6 into 2
  
# split part of class 1 into 8
clusters <- within(clusters, class2[class2 == 1 & magnitude_homerange == -2] <- 8) 
clusters <- within(clusters, class2[class2 == 1 & magnitude_homerange == -3] <- 8)









```

```{r}
fish <- load_taxa() %>%  # load all taxonomic info from fishbase into fish, then
  filter(Species %in% data$species) # select only the rows containing species listed in data
```


### Collect diet variable

```{r}
# get diet items and combine rows of the same species in each column
diet <- fooditems(fish$Species, fields = c("Species", "FoodI", "FoodII", "FoodIII")) %>% 
  group_by(Species) %>%
  summarise(FoodI = paste(unique(FoodI), collapse = ", "), 
            FoodII = paste(unique(FoodII), collapse = ", "), 
            FoodIII = paste(unique(FoodIII), collapse = ", ")) %>% 
  ungroup()

# using unique(x) makes sure there are not duplicates

diet <- janitor::clean_names(diet)

data <- left_join(data,diet) %>% 
  select(-food_ii, -food_iii)
```


### Collect location variables

```{r}
# get location data (lat and long)
coords <- stocks() %>%  # load stocks info, then
  filter(SpecCode %in% fish$SpecCode) %>%  # filter rows by specCodes in fish, then
  select(SpecCode, Northernmost, NorthSouthN, Southermost, NorthSouthS, Westernmost,
           WestEastW, Easternmost, WestEastE) # select columns

coords <- na.omit(coords) # remove NA values

# not sure if I need to make everything uppercase?

coords <- data.frame(lapply(coords, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
}))

# there are a few spec codes listed more than once. 
# can I delete the duplicated rows..? there are only 5 <-----------------------------------------------

coords_clean <- coords[!duplicated(coords$SpecCode), ]

# add coords to fish because they share a column
fish <- left_join(fish, coords_clean)

fish <- janitor::clean_names(fish)
```


### Collect temperature variables

```{r}

temp <- stocks() %>%  # load stocks info, then
  filter(SpecCode %in% fish$spec_code) %>%  # filter rows by spec codes in fish, then
  select(SpecCode, TempMin, TempMax, TempPreferred, EnvTemp) %>% # select columns
  mutate(temp_range = TempMax - TempMin)

temp <- drop_na(temp, TempPreferred)

temp <- janitor::clean_names(temp)


fish <- left_join(fish, temp)

```


### Collect fecundity variables

```{r}

fecund <- fecundity() %>%  # load stocks info, then
  filter(SpecCode %in% fish$spec_code) %>%  # filter rows by spec codes in fish, then
  select(SpecCode, FecundityMin, FecundityMax) %>%   # select columns
  janitor::clean_names() %>% 
  na.omit(fecund)

fecund <- fecund %>%  # average the values for a given spec code
  group_by(spec_code) %>% 
  summarise(fecundity_min = mean(fecundity_min), fecundity_max = mean(fecundity_max)) %>% 
  mutate(fecundity_mean = (fecundity_min + fecundity_max)/2) %>% 
  ungroup()


fish <- left_join(fish, fecund)

```


### Collect shape/age data

```{r}

spec <- species() %>%  # load stocks info, then
  filter(SpecCode %in% fish$spec_code) %>%  # filter rows by spec codes in fish, then
  select(SpecCode, BodyShapeI, LongevityWild) %>% 
  janitor::clean_names()


fish <- left_join(fish, spec)

```


### Collect metabolic data

Only 68 species with oxygen consumption data:
```{r}

oxygen <- oxygen() %>% 
  filter(SpecCode %in% fish$spec_code) %>% 
  select(SpecCode, OxygenCons, Temp) %>% 
  janitor::clean_names()

oxygen <- oxygen %>%  # average the values for a given spec code
  group_by(spec_code) %>% 
  summarise(oxygen_cons = mean(oxygen_cons), temp = mean(temp)) %>% 
  ungroup() %>% 
  rename(temp_cons = temp)

fish <- left_join(fish, oxygen)
fish <- fish %>% 
  rename(class_taxon = class)


```


### Join new variables with data

```{r}

# why did this add a row? <------------------------------------------------------------------------------

data <- left_join(fish, data) %>% 
  select(-spec_code, -class, -fecundity_min, -temp_max, -temp_min)

data <- full_join(data, clusters)


# add missing info

data$body_shape_i[data$species == "Oblada melanuras"] <- "fusiform / normal"
data$body_shape_i[data$species == "Pennahia aneas"] <- "fusiform / normal"

# ADD MORE INFO ON THESE TWO GUYS HERE

```


### Write new csv including added variables

```{r}

#write_csv(data, here::here("01_data", "02_processed_data", "fishbase_data.csv"))

```


### fill in missing info from averages/max counts

```{r}

# create a function to find the mode of a column

calc_mode <- function(x){  
  
  distinct_values <- unique(x) # list the unique values
  
  distinct_tabulate <- tabulate(match(x, distinct_values)) # count the occurrence of each
  
  distinct_values[which.max(distinct_tabulate)] # return value with highest occurrence
}


# replace numeric NAs with mean of column, categorical with mode of column

filled_data <- data %>%  
  mutate_at(vars(temp_preferred, longevity_wild, fecundity_max, fecundity_mean, 
                 temp_range, temp_cons, oxygen_cons), 
            ~replace_na(.,mean(., na.rm = TRUE))) %>% 
  mutate(food_i = if_else(is.na(food_i), calc_mode(food_i), food_i)) %>% 
  mutate(env_temp = if_else(is.na(env_temp), calc_mode(env_temp), env_temp)) %>% 
  mutate(iucn_category = if_else(is.na(iucn_category), calc_mode(iucn_category), iucn_category))

```


### Write new csv including filled variables

```{r}

#write_csv(filled_data, here::here("01_data", "02_processed_data", "filled_data.csv"))

```


### Look at missing data

```{r}

missing_data <- miss_var_summary(data) # summary of missing data in each variable
  
missing <- data %>% 
  select(- species, -genus, -subfamily, -family, -order, -class_taxon, -super_class, -body_shape_i, -temp_cons, -n_ind, -geog_range, -fecundity_mean, -movement_keyword, -demers_pelag, -length, -common_name, -kfin, -r_fin, -diet_troph3, -iucn_category, -predicted_homerange, -predicted_pld, -home_range, -pld, -magnitude_homerange, -month_pld, -class, -class2, -north_south_n, -southermost, -north_south_s, -westernmost, -west_east_w, -easternmost, -west_east_e) %>% 
  rename(geographic_coords = northernmost)
  
missing_fig <- gg_miss_var(missing) + # figure showing the missing data
  theme_bw()
missing_fig

#ggsave(missing_fig, file = paste0("missing_fig.png"), path = here::here("04_figs"))


```


### Look at any correlated data

```{r}

# for numeric variables, omit NAs?
data_cor <- data %>% 
  select(where(is.numeric)) %>% 
  select(length, kfin, r_fin, diet_troph3, temp_preferred, 
         fecundity_mean, longevity_wild, weight, age_mat_min, age_mat_min2) %>% 
  na.omit()

cor_mat <- round(cor(data_cor),2) # length and weight = 0.88, age_mat_min/2 and longevity_wild > 0.70 for all three
cor_mat


```









### Plot of new clusters

```{r}

fishbase_data_gmm <- data %>% 
  mutate(class2 = as.factor(class2))

gmm_new_fig <- ggplot(fishbase_data_gmm, aes(magnitude_homerange, month_pld)) +
  geom_jitter(aes(color = class2)) +
  theme_bw() +
  labs(
    x = "Order of Magnitude of Predicted Home Range",
    y = "Predicted Pelagic Larval Duration (Month)"
  )

gmm_new_fig

# ggsave(gmm_new_fig, file = paste0("gmm_new_fig.png"), path = here::here("04_figs"), height = 8, width = 15)

```







### Old

```{r}

# collect maturity variables

# matur <- maturity() %>% 
#   filter(SpecCode %in% fish$spec_code) %>% 
#   select(SpecCode, AgeMatMin, AgeMatMin2) %>% 
#   janitor::clean_names() %>% 
#   na.omit()
# 
# matur <- matur %>%  # average the values for a given spec code
#   group_by(spec_code) %>% 
#   summarise(age_mat_min = mean(age_mat_min), age_mat_min2 = mean(age_mat_min2)) %>% 
#   ungroup()
# 
# fish <- left_join(fish, matur)
```

```{r}
# 
# # Only 14 species with schooling behavior listed:
# 
# eco <- ecology() %>% 
#   filter(SpecCode %in% fish$spec_code) %>% 
#   select(SpecCode, SchoolingFrequency, SchoolingLifestage) %>% 
#   janitor::clean_names()

```

```{r}
# can I average the lat/long values and delete the duplicated rows? <--------------- NO. bc this messes with the E/W delineation

# mean_stocks <- stocks %>% 
  # group_by(SpecCode) %>% 
  # summarise(Northernmost = mean(Northernmost), NorthSouthN = NorthSouthN, Southernmost =
             # mean(Southermost),NorthSouthS = NorthSouthS, Westernmost = mean(Westernmost), WestEastW =
             # WestEastW, Easternmost = mean(Easternmost), WestEastE = WestEastE) %>% 
  # ungroup()
```

```{r}
# created a dataframe with 3 fewer rows?

# order(data$species) # order the dataframes by species to look for missing species names
# order(fish$Species)
# data$species[!(data$species %in% fish$Species)] # find the species in data that are not included in fish

# 3 species not included in fish: Limanda ferruginea, Oblada melanura, Pennahia anea

# Myzopsetta ferruginea for Limanda ferruginea, Oblada melanurus for Oblada melanura, Pennahia aneas for Pennahia anea
```



