---
title: "PCA"
output: html_document
date: "2023-04-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(rfishbase)
library(FactoMineR)
library(factoextra)
#library(vcd)
#library(purrr)
library(car)
library(vegan)
library(patchwork)
library(cowplot)
library(ggpubr)
library(naniar)
#library(microViz) # for betadisp categorical wrapper
#library(phyloseq) # for betadisp categorical wrapper
```

### Load data

```{r}
filled_data2 <- read_csv(here::here("01_data", "02_processed_data", "filled_data2.csv")) %>%
  mutate_if(is.character, as.factor)
```


### Plot data

```{r}
# continuous

plot_data <- filled_data2 %>%
  mutate(class2 = as.factor(class2), 
         class3 = as.factor(class3), 
         order = as.factor(order)
         )

response <- "class2"
expl_cont <- c(
  "temp_preferred", "fecundity_mean",
  "longevity_wild", "sp_richness", "length", "kfin",
  "r_fin", "diet_troph3", "oxygen_cons", "temp_range"
)

box_data <- function(x, y) {
  ggplot(plot_data, aes(x = .data[[y]], y = .data[[x]])) +
    geom_boxplot(aes(color = .data[[y]])) +
    theme_bw() +
    coord_flip() +
    labs(
      x = "Movement Class",
      color = "Class (homerange/PLD)"
    ) +
    scale_color_viridis_d()
}

#plots_cont <- map(expl_cont, ~ box_data(.x, "class2"))
#plots_cont
```

```{r}
temp_plot <- box_data(x = "temp_preferred", y = "class2") + stat_kruskal_test(label.y = 5, position = "jitter")
fecund_plot <- box_data(x = "fecundity_mean", y = "class2") + stat_kruskal_test(label.y = 100000, position = "jitter")
long_plot <- box_data(x = "longevity_wild", y = "class2") + stat_kruskal_test(label.y = 50, position = "jitter")
sp_rich_plot <- box_data(x = "sp_richness", y = "class2") + stat_kruskal_test(label.y = 500, position = "jitter")
length_plot <- box_data(x = "length", y = "class2") + stat_anova_test(label.y = 200, position = "jitter")
kfin_plot <- box_data(x = "kfin", y = "class2") + stat_kruskal_test(label.y = 10000000, position = "jitter")
r_fin_plot <- box_data(x = "r_fin", y = "class2") + stat_kruskal_test(label.y = 1, position = "jitter")
diet_plot <- box_data(x = "diet_troph3", y = "class2") + stat_kruskal_test(label.y = 2.5, position = "jitter")
oxygen_plot <- box_data(x = "oxygen_cons", y = "class2") + stat_kruskal_test(label.y = 2000, position = "jitter")
temp_range_plot <- box_data(x = "temp_range", y = "class2") + stat_kruskal_test(label.y = 15, position = "jitter")


ggsave(temp_plot, file = paste0("temp_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10) 
ggsave(fecund_plot, file = paste0("fecund_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(long_plot, file = paste0("long_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(sp_rich_plot, file = paste0("sp_rich_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(length_plot, file = paste0("length_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(kfin_plot, file = paste0("kfin_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(r_fin_plot, file = paste0("r_fin_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(diet_plot, file = paste0("diet_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(oxygen_plot, file = paste0("oxygen_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)
ggsave(temp_range_plot, file = paste0("temp_range_plot.png"), path = here::here("04_figs", "individual_plots", "numeric"), height = 10, width = 10)

```

```{r}
# categorical

expl_cat <- c(
  "climate", "body_shape_i", "dev_place", "movement_keyword",
  "demers_pelag", "iucn_category",
  "circadian", "food_mode", "ocean", "basin"
)

scatter_data <- function(x, y) {
  ggplot(plot_data, aes(x = .data[[x]], y = .data[[y]])) +
    geom_jitter(aes(color = .data[[y]])) +
    theme_bw() +
    labs(
      y = "Movement Class",
      color = "Class (homerange/PLD)"
    ) +
    scale_color_viridis_d() +
    theme(axis.text.x = element_text(angle = 40, hjust = 1))
}

#plots_cat <- map(expl_cat, ~ scatter_data(.x, "class2"))
#plots_cat
```

```{r}
climate_plot <- scatter_data(x = "climate", y = "class2") # + stat_kruskal_test(label.y = 5, label.x = "6: high/mid")
body_plot <- scatter_data(x = "body_shape_i", y = "class2")
dev_plot <- scatter_data(x = "dev_place", y = "class2")
movement_plot <- scatter_data(x = "movement_keyword", y = "class2")
demers_plot <- scatter_data(x = "demers_pelag", y = "class2")
iucn_plot <- scatter_data(x = "iucn_category", y = "class2")
food_plot <- scatter_data(x = "food_mode", y = "class2")
# circ_plot <- scatter_data(x = "circadian", y = "class2")
# ocean_plot <- scatter_data(x = "ocean", y = "class2")
# basin_plot <- scatter_data(x = "basin", y = "class2")


ggsave(climate_plot, file = paste0("climate_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
ggsave(body_plot, file = paste0("body_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
ggsave(dev_plot, file = paste0("dev_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
ggsave(movement_plot, file = paste0("movement_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
ggsave(demers_plot, file = paste0("demers_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
ggsave(iucn_plot, file = paste0("iucn_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
ggsave(food_plot, file = paste0("food_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
# ggsave(circ_plot, file = paste0("circ_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
# ggsave(ocean_plot, file = paste0("ocean_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)
# ggsave(basin_plot, file = paste0("basin_plot.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 10, width = 10)

```


### Plot panel

```{r}

cat_first_col <- (demers_plot / body_plot / climate_plot) &
  theme(legend.position = "none")

cat_second_col <- (movement_plot / dev_plot / food_plot) &
  theme(legend.position = "none") &
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())

cat_top <- (cat_first_col | cat_second_col)

cat_bottom <- (iucn_plot | plot_spacer()) + plot_layout(widths = c(55,30))

cat_panel <- (cat_top / cat_bottom) + plot_layout(heights = c(85,20))

cat_panel

ggsave(cat_panel, file = paste0("cat_panel.png"), path = here::here("04_figs", "individual_plots", "categorical"), height = 14, width = 10)

```




### Check Variable Inflation Factors (VIF)

none are above 5

```{r}
model <- lm(
  class3 ~ temp_preferred+ fecundity_mean+ longevity_wild+
    sp_richness+ length+ kfin+ r_fin+ diet_troph3,
  data = filled_data2
)

summary(model)

vif(model) # when including both fecundity variables, they have a VIF over 5

# create vector of VIF values
vif_values <- vif(model)

# create horizontal bar chart to display each VIF value
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue", xlim = c(0, 6), las = 2)

# add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)


# model2 <- lm(class2 ~ temp_preferred + length + r_fin, data = filled_data)
# summary(model)
```


### Similarity percentage

```{r}

com_data <- filled_data2 %>%
  dplyr::select(
    temp_preferred, fecundity_mean, longevity_wild,
    sp_richness, length, kfin, r_fin, diet_troph3, species, family, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>% 
  filter(temp_preferred > 0) %>%
  mutate(across(where(is.numeric),  ~ scale(.)[,1])) %>% 
  na.omit()

com <- com_data %>%
  dplyr::select(-species, -family, -class2)

# load from rda in results folder
#set.seed(123)
#simper_out = simper(comm = com, group = com_data$class2, permutations = 9999)

summary(simper_out)
# shows which traits are primarily responsible for observed difference between classes. traits listed from highest to lowest % contribution to the difference between the two classes (cumsum shows total % contribution to the difference). p is signifigance.


# load from rda in results folder
#set.seed(123)
#ano = anosim(com, com_data$class2, distance = "euclidean", permutations = 9999)

summary(ano)

plot(ano)
```


### FAMD

```{r}

famd_1 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, fecundity_mean, longevity_wild, sp_richness, length, 
    kfin, r_fin, diet_troph3, oxygen_cons, temp_range, climate, body_shape_i, 
    dev_place, iucn_category, food_mode, demers_pelag,
    class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()


f1 <- FAMD(famd_1[, 1:16], graph = FALSE)
f1

# Contribution to the first dimension
fviz_contrib(f1, "var", axes = 1) + fviz_contrib(f1, "var", axes = 2)
# Contribution to the second dimension


# PLOTS ARROWS
p1 <- fviz_famd_var(f1, "quanti.var",
  repel = TRUE,
  col.var = "black"
)

eigen_f1 <- f1$eig
eigen_f1


f1_fig <- fviz_famd_ind(f1, label = "none", habillage = famd_1$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f1_fig

plot1 <- p1 + f1_fig

plot1

#ggsave(plot1, file = paste0("f1.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```


```{r}
famd_2 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, longevity_wild, body_shape_i, length,
    r_fin, diet_troph3, temp_range, climate, 
    demers_pelag, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()


f2 <- FAMD(famd_2[, 1:9], graph = FALSE)
f2

eigen_f2 <- f2$eig
eigen_f2

# Contribution to the first dimension
fviz_contrib(f2, "var", axes = 1) + fviz_contrib(f2, "var", axes = 2)
# Contribution to the second dimension



# PLOTS ARROWS
p2 <- fviz_famd_var(f2, "quanti.var",
  repel = TRUE,
  col.var = "black"
)



f2_fig <- fviz_famd_ind(f2, label = "none", habillage = famd_2$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f2_fig

plot2 <- p2 + f2_fig

plot2

#ggsave(plot2, file = paste0("f2.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)
```




```{r}
famd_4 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, length, r_fin, dev_place, demers_pelag, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()

f4 <- FAMD(famd_4[, 1:5], graph = FALSE)
f4

eigen_f4 <- f4$eig
eigen_f4

# Contribution to the first dimension
fviz_contrib(f4, "var", axes = 1) + fviz_contrib(f4, "var", axes = 2)
# Contribution to the second dimension



# PLOTS ARROWS
p4 <- fviz_famd_var(f4, "quanti.var",
  repel = TRUE,
  col.var = "black"
)


f4_fig <- fviz_famd_ind(f4, label = "none", habillage = famd_4$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f4_fig

plot4 <- p4 + f4_fig

plot4

#ggsave(plot4, file = paste0("f4-1.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)
```


```{r}

famd_7 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, length, demers_pelag, oxygen_cons, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()

f7 <- FAMD(famd_7[,1:4], graph = FALSE)
f7

eigen_f7 <- f7$eig
eigen_f7

# Contribution to the first dimension
fviz_contrib(f7, "var", axes = 1) + fviz_contrib(f7, "var", axes = 2)
# Contribution to the second dimension



# PLOTS ARROWS
p7<- fviz_famd_var(f7, "quanti.var",
  repel = TRUE,
  col.var = "black")


f7_fig <- fviz_famd_ind(f7, label = "none", habillage = famd_7$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f7_fig

plot7 <- p7 + f7_fig

plot7

#ggsave(plot7, file = paste0("f7.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```


```{r}

famd_10 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, length, demers_pelag, diet_troph3, longevity_wild, sp_richness, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()

f10 <- FAMD(famd_10[,1:6], graph = FALSE)
f10

eigen_f10 <- f10$eig
eigen_f10

# Contribution to the first dimension
fviz_contrib(f10, "var", axes = 1) + fviz_contrib(f10, "var", axes = 2)
# Contribution to the second dimension



# PLOTS ARROWS
p10<- fviz_famd_var(f10, "quanti.var",
  repel = TRUE,
  col.var = "black")


f10_fig <- fviz_famd_ind(f10, label = "none", habillage = famd_10$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f10_fig

plot10 <- p10 + f10_fig

plot10

#ggsave(plot10, file = paste0("f10.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```


```{r}

famd_12 <- filled_data2 %>%
  dplyr::select(
    length, demers_pelag, longevity_wild, r_fin, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()

f12 <- FAMD(famd_12[,1:4], graph = FALSE)
f12

eigen_f12 <- f12$eig
eigen_f12

# Contribution to the first dimension
fviz_contrib(f12, "var", axes = 1) + fviz_contrib(f12, "var", axes = 2)
# Contribution to the second dimension



# PLOTS ARROWS
p12<- fviz_famd_var(f12, "quanti.var",
  repel = TRUE,
  col.var = "black")


f12_fig <- fviz_famd_ind(f12, label = "none", habillage = famd_12$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f12_fig

plot12 <- p12 + f12_fig

plot12

#ggsave(plot12, file = paste0("f12.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```


```{r}

famd_13 <- filled_data2 %>%
  dplyr::select(
    demers_pelag, longevity_wild, temp_range, temp_preferred, class2
  ) %>%
  mutate(class2 = as.factor(class2)) %>%
  na.omit()

f13 <- FAMD(famd_13[,1:4], graph = FALSE)
f13

eigen_f13 <- f13$eig
eigen_f13

# Contribution to the first dimension
fviz_contrib(f13, "var", axes = 1) + fviz_contrib(f13, "var", axes = 2)
# Contribution to the second dimension



# PLOTS ARROWS
p13<- fviz_famd_var(f13, "quanti.var",
  repel = TRUE,
  col.var = "black")


f13_fig <- fviz_famd_ind(f13, label = "none", habillage = famd_13$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
  theme_bw()

f13_fig

plot13 <- p13 + f13_fig

plot13

#ggsave(plot13, file = paste0("f13.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```






### NMDS 1

```{r}
com_data <- filled_data2 %>%
  dplyr::select(
    temp_preferred, fecundity_mean, longevity_wild,
    sp_richness, length, kfin, r_fin, diet_troph3, oxygen_cons, 
    temp_range, species, order, class2
  ) %>%
  filter(temp_preferred > 0) %>%
  na.omit()

com <- com_data %>%
  dplyr::select(-species, -order, -class2)

m_com <- as.matrix(com) 

set.seed(123)
nmds <- metaMDS(m_com, distance = "euclidean")
nmds

# extract NMDS scores (x and y coordinates)
filled_scores <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores$class2 <- com_data$class2
filled_scores$species <- com_data$species
filled_scores$order <- com_data$order

filled_scores <- filled_scores %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")

centroid <- filled_scores %>% 
  group_by(class2) %>% 
  summarize(NMDS1 = mean(NMDS1), NMDS2 = mean(NMDS2))

nmds_plot1 <- ggplot(filled_scores, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  geom_point(data = centroid, aes(color = class2), size = 4, show.legend = FALSE) +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.15, show.legend = FALSE) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)


nmds_plot1

#ggsave(nmds_plot1, file = paste0("nmds_1.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst <- dist(com, method = "euclidean")

bd <- betadisper(dst, com_data$class2)

anova(bd)

permutest(bd, pairwise = TRUE, permutations = 9999)

#bd_HSD <- TukeyHSD(bd)
#plot(bd_HSD)
#
# labs <- paste("Dimension", 1:10, "(", 
#               round(100*bd$eig / sum(bd$eig), 2), "%)")
# plot(bd,
#      main="MDS coordinates",
#      xlab=labs[1], ylab=labs[2],
#      hull=FALSE, ellipse=TRUE, conf=0.95, lwd=2, label = FALSE)

```


##### NMDS 2

```{r}
com_data2 <- filled_data2 %>%
  dplyr::select(
    longevity_wild,
    sp_richness, kfin, r_fin, oxygen_cons, 
    temp_range, species, order, class2
  ) %>%
  na.omit()

com2 <- com_data2 %>%
  dplyr::select(-species, -order, -class2)

m_com2 <- as.matrix(com2) 

set.seed(123)
nmds <- metaMDS(m_com2, distance = "euclidean")
nmds 

# extract NMDS scores (x and y coordinates)
filled_scores2 <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores2$class2 <- com_data2$class2
filled_scores2$species <- com_data2$species
filled_scores2$order <- com_data2$order

filled_scores2 <- filled_scores2 %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")

centroid2 <- filled_scores2 %>% 
  group_by(class2) %>% 
  summarize(NMDS1 = mean(NMDS1), NMDS2 = mean(NMDS2))

nmds_plot2 <- ggplot(filled_scores2, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  geom_point(data = centroid2, aes(color = class2), size = 4, show.legend = FALSE) +
  # theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.15, show.legend = FALSE) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot2

#ggsave(nmds_plot2, file = paste0("nmds_2.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst2 <- dist(com2, method = "euclidean")

bd2 <- betadisper(dst2, com_data2$class2)

anova(bd2)

permutest(bd2, pairwise = TRUE, permutations = 9999)

```



##### NMDS 5

```{r}
com_data5 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, longevity_wild, length, r_fin, diet_troph3, temp_range,
    species, order, class2
  ) %>%
  mutate(temp_preferred = temp_preferred + 2) %>%
  na.omit()

com5 <- com_data5 %>%
  dplyr::select(-species, -order, -class2)

m_com5 <- as.matrix(com5) 

set.seed(123)
nmds <- metaMDS(m_com5, distance = "euclidean")
nmds


# extract NMDS scores (x and y coordinates)
filled_scores5 <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores5$class2 <- com_data5$class2
filled_scores5$species <- com_data5$species
filled_scores5$order <- com_data5$order

filled_scores5 <- filled_scores5 %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")


nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot5

#ggsave(nmds_plot5, file = paste0("nmds_5.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst5 <- dist(com5, method = "euclidean")

bd5 <- betadisper(dst5, com_data5$class2)

anova(bd5)

permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS 12

```{r}
com_data12 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, r_fin, diet_troph3, temp_range,
    species, order, class2
  ) %>%
  mutate(temp_preferred = temp_preferred + 2) %>%
  na.omit()

com12 <- com_data12 %>%
  dplyr::select(-species, -order, -class2)

m_com12 <- as.matrix(com12) 

set.seed(123)
nmds <- metaMDS(m_com12, distance = "euclidean")
nmds


# extract NMDS scores (x and y coordinates)
filled_scores12 <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores12$class2 <- com_data12$class2
filled_scores12$species <- com_data12$species
filled_scores12$order <- com_data12$order

filled_scores12 <- filled_scores12 %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")


nmds_plot12 <- ggplot(filled_scores12, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot12

#ggsave(nmds_plot12, file = paste0("nmds_12.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst12 <- dist(com12, method = "euclidean")

bd12 <- betadisper(dst12, com_data12$class2)

anova(bd12)

permutest(bd12, pairwise = TRUE, permutations = 9999)

```


##### NMDS Plot 13

```{r}
com_data13 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, r_fin, diet_troph3, sp_richness, temp_range,
    species, order, class2
  ) %>%
  mutate(temp_preferred = temp_preferred + 2) %>%
  na.omit()

com13 <- com_data13 %>%
  dplyr::select(-species, -order, -class2)

m_com13 <- as.matrix(com13)

set.seed(123)
nmds <- metaMDS(m_com13, distance = "euclidean")
nmds

# extract NMDS scores (x and y coordinates)
filled_scores13 <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores13$class2 <- com_data13$class2
filled_scores13$species <- com_data13$species
filled_scores13$order <- com_data13$order

filled_scores13 <- filled_scores13 %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")


nmds_plot13 <- ggplot(filled_scores13, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot13

#ggsave(nmds_plot13, file = paste0("nmds_13.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst13 <- dist(com13, method = "euclidean")

bd13 <- betadisper(dst13, com_data13$class2)

anova(bd13)

permutest(bd13, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 14

```{r}
com_data14 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, r_fin, diet_troph3, sp_richness, kfin, temp_range,
    species, order, class2
  ) %>%
  mutate(temp_preferred = temp_preferred + 2) %>%
  na.omit()

com14 <- com_data14 %>%
  dplyr::select(-species, -order, -class2)

m_com14 <- as.matrix(com14) 

set.seed(123)
nmds <- metaMDS(m_com14, distance = "euclidean")
nmds


# extract NMDS scores (x and y coordinates)
filled_scores14 <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores14$class2 <- com_data14$class2
filled_scores14$species <- com_data14$species
filled_scores14$order <- com_data14$order

filled_scores14 <- filled_scores14 %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")


nmds_plot14 <- ggplot(filled_scores14, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot14

#ggsave(nmds_plot14, file = paste0("nmds_14.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst14 <- dist(com14, method = "euclidean")

bd14 <- betadisper(dst14, com_data14$class2)

anova(bd14)

permutest(bd14, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 17

```{r}
com_data17 <- filled_data2 %>%
  dplyr::select(
    temp_preferred, r_fin, diet_troph3, oxygen_cons,
    species, order, class2
  ) %>%
  mutate(temp_preferred = temp_preferred + 2) %>%
  na.omit()

com17 <- com_data17 %>%
  dplyr::select(-species, -order, -class2)

m_com17 <- as.matrix(com17) 

set.seed(123)
nmds <- metaMDS(m_com17, distance = "euclidean")
nmds

# extract NMDS scores (x and y coordinates)
filled_scores17 <- as.data.frame(scores(nmds)$sites)

# add columns to data frame
filled_scores17$class2 <- com_data17$class2
filled_scores17$species <- com_data17$species
filled_scores17$order <- com_data17$order

filled_scores17 <- filled_scores17 %>%
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores <- as.data.frame(nmds[["species"]]) %>%
  rownames_to_column(var = "var")


nmds_plot17 <- ggplot(filled_scores17, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
  geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot17

#ggsave(nmds_plot17, file = paste0("nmds_17.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)

dst17 <- dist(com17, method = "euclidean")

bd17 <- betadisper(dst17, com_data17$class2)

anova(bd17)

permutest(bd17, pairwise = TRUE, permutations = 9999)

```
















# Old


```{r}
# famd_3 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, r_fin, diet_troph3, climate, 
#     demers_pelag, class2
#   ) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
# 
# 
# f3 <- FAMD(famd_3[, 1:6], graph = FALSE)
# f3
# 
# eigen_f3 <- f3$eig
# eigen_f3
# 
# # Contribution to the first dimension
# fviz_contrib(f3, "var", axes = 1) + fviz_contrib(f3, "var", axes = 2)
# # Contribution to the second dimension
# 
# 
# 
# # PLOTS ARROWS
# p3 <- fviz_famd_var(f3, "quanti.var",
#   repel = TRUE,
#   col.var = "black"
# )
# 
# 
# 
# f3_fig <- fviz_famd_ind(f3, label = "none", habillage = famd_3$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
#   theme_bw()
# 
# f3_fig 
# 
# plot3 <- p3 + f3_fig
# 
# plot3
# 
# #ggsave(plot3, file = paste0("f3.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)
```


```{r}
# famd_5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, kfin, dev_place, demers_pelag, class2
#   ) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
# 
# f5 <- FAMD(famd_5[,1:5], graph = FALSE)
# f5
# 
# eigen_f5 <- f5$eig
# eigen_f5
# 
# # Contribution to the first dimension
# fviz_contrib(f5, "var", axes = 1) + fviz_contrib(f5, "var", axes = 2)
# # Contribution to the second dimension
# 
# 
# 
# # PLOTS ARROWS
# p5<- fviz_famd_var(f5, "quanti.var",
#   repel = TRUE,
#   col.var = "black")
# 
# 
# f5_fig <- fviz_famd_ind(f5, label = "none", habillage = famd_5$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
#   theme_bw()
# 
# f5_fig
# 
# plot5 <- p5 + f5_fig
# 
# plot5
# 
# #ggsave(plot5, file = paste0("f5.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)
```


```{r}
# 
# famd_6 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, climate, demers_pelag, class2
#   ) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
# 
# f6 <- FAMD(famd_6[,1:4], graph = FALSE)
# f6
# 
# eigen_f6 <- f6$eig
# eigen_f6
# 
# # Contribution to the first dimension
# fviz_contrib(f6, "var", axes = 1) + fviz_contrib(f6, "var", axes = 2)
# # Contribution to the second dimension
# 
# 
# 
# # PLOTS ARROWS
# p6<- fviz_famd_var(f6, "quanti.var",
#   repel = TRUE,
#   col.var = "black")
# 
# 
# f6_fig <- fviz_famd_ind(f6, label = "none", habillage = famd_6$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
#   theme_bw()
# 
# f6_fig
# 
# plot6 <- p6 + f6_fig
# 
# plot6
# 
# #ggsave(plot6, file = paste0("f6.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```

```{r}

# famd_8 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, demers_pelag, kfin, class2
#   ) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
# 
# f8 <- FAMD(famd_8[,1:4], graph = FALSE)
# f8
# 
# eigen_f8 <- f8$eig
# eigen_f8
# 
# # Contribution to the first dimension
# fviz_contrib(f8, "var", axes = 1) + fviz_contrib(f8, "var", axes = 2)
# # Contribution to the second dimension
# 
# 
# 
# # PLOTS ARROWS
# p8<- fviz_famd_var(f8, "quanti.var",
#   repel = TRUE,
#   col.var = "black")
# 
# 
# f8_fig <- fviz_famd_ind(f8, label = "none", habillage = famd_8$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
#   theme_bw()
# 
# f8_fig
# 
# plot8 <- p8 + f8_fig
# 
# plot8
# 
# #ggsave(plot8, file = paste0("f8.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```


```{r}

# famd_9 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, demers_pelag, diet_troph3, class2
#   ) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
# 
# f9 <- FAMD(famd_9[,1:4], graph = FALSE)
# f9
# 
# eigen_f9 <- f9$eig
# eigen_f9
# 
# # Contribution to the first dimension
# fviz_contrib(f9, "var", axes = 1) + fviz_contrib(f9, "var", axes = 2)
# # Contribution to the second dimension
# 
# 
# 
# # PLOTS ARROWS
# p9<- fviz_famd_var(f9, "quanti.var",
#   repel = TRUE,
#   col.var = "black")
# 
# 
# f9_fig <- fviz_famd_ind(f9, label = "none", habillage = famd_9$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
#   theme_bw()
# 
# f9_fig
# 
# plot9 <- p9 + f9_fig
# 
# plot9
# 
# #ggsave(plot9, file = paste0("f9.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)

```

```{r}

# famd_11 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, demers_pelag, diet_troph3, longevity_wild, sp_richness, kfin, class2
#   ) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
# 
# f11 <- FAMD(famd_11[,1:7], graph = FALSE)
# f11
# 
# eigen_f11 <- f11$eig
# eigen_f11
# 
# # Contribution to the first dimension
# fviz_contrib(f11, "var", axes = 1) + fviz_contrib(f11, "var", axes = 2)
# # Contribution to the second dimension
# 
# 
# 
# # PLOTS ARROWS
# p11<- fviz_famd_var(f11, "quanti.var",
#   repel = TRUE,
#   col.var = "black")
# 
# 
# f11_fig <- fviz_famd_ind(f11, label = "none", habillage = famd_11$class2, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.var") +
#   theme_bw()
# 
# f11_fig
# 
# plot11 <- p11 + f11_fig
# 
# plot11
# 
# #ggsave(plot11, file = paste0("f11.png"), path = here::here("04_figs", "famd_with_arrows"), height = 12, width = 15)


```

##### NMDS Plot 3

```{r}
# com_data3 <- filled_data2 %>%
#   dplyr::select(
#     longevity_wild,
#     sp_richness, kfin, oxygen_cons, 
#     species, order, class2
#   ) %>%
#   na.omit()
# 
# com3 <- com_data3 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com3 <- as.matrix(com3) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com3, distance = "euclidean")
# nmds 
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores3 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores3$class2 <- com_data3$class2
# filled_scores3$species <- com_data3$species
# filled_scores3$order <- com_data3$order
# 
# filled_scores3 <- filled_scores3 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot3 <- ggplot(filled_scores3, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot3
# 
# #ggsave(nmds_plot3, file = paste0("nmds_3.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst3 <- dist(com3, method = "euclidean")
# 
# bd3 <- betadisper(dst3, com_data3$class2)
# 
# anova(bd3)
# 
# permutest(bd3, pairwise = TRUE, permutations = 9999)

```

##### NMDS Plot 4

```{r}
# com_data4 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, longevity_wild, sp_richness, 
#     length, kfin, r_fin, diet_troph3,
#     species, order, class2
#   ) %>%
#   filter(temp_preferred > 0) %>%
#   na.omit()
# 
# com4 <- com_data4 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com4 <- as.matrix(com4) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com4, distance = "euclidean")
# nmds
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores4 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores4$class2 <- com_data4$class2
# filled_scores4$species <- com_data4$species
# filled_scores4$order <- com_data4$order
# 
# filled_scores4 <- filled_scores4 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# nmds_plot4 <- ggplot(filled_scores4, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot4
# 
# #ggsave(nmds_plot4, file = paste0("nmds_4.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst4 <- dist(com4, method = "euclidean")
# 
# bd4 <- betadisper(dst4, com_data4$class2)
# 
# anova(bd4)
# 
# permutest(bd4, pairwise = TRUE, permutations = 9999)

```

##### NMDS Plot 6

```{r}
# com_data6 <- filled_data2 %>%
#   dplyr::select(
#     sp_richness, kfin, oxygen_cons, length,
#     species, order, class2
#   ) %>%
#   na.omit()
# 
# com6 <- com_data6 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com6 <- as.matrix(com6) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com6, distance = "euclidean")
# nmds
# 
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores6 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores6$class2 <- com_data6$class2
# filled_scores6$species <- com_data6$species
# filled_scores6$order <- com_data6$order
# 
# filled_scores6 <- filled_scores6 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# nmds_plot6 <- ggplot(filled_scores6, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot6
# 
# #ggsave(nmds_plot6, file = paste0("nmds_6.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst6 <- dist(com6, method = "euclidean")
# 
# bd6 <- betadisper(dst6, com_data6$class2)
# 
# anova(bd6)
# 
# permutest(bd6, pairwise = TRUE, permutations = 9999)

```


##### NMDS Plot 7

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, longevity_wild, length, r_fin, sp_richness, temp_range,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5)
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_7.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 8

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, r_fin, diet_troph3, temp_range,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_8.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 9

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, r_fin, diet_troph3, temp_range,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_9.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 10

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, diet_troph3, temp_range,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_10.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 11

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, longevity_wild, diet_troph3, temp_range,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_11.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```

##### NMDS Plot 15

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, r_fin, diet_troph3, longevity_wild,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_15.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 16

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, r_fin, diet_troph3, length,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_16.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```


##### NMDS Plot 18

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, r_fin, diet_troph3, fecundity_mean,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_18.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```



##### NMDS Plot 19

```{r}
# com_data5 <- filled_data2 %>%
#   dplyr::select(
#     temp_preferred, length, diet_troph3, fecundity_mean,
#     species, order, class2
#   ) %>%
#   mutate(temp_preferred = temp_preferred + 2) %>%
#   na.omit()
# 
# com5 <- com_data5 %>%
#   dplyr::select(-species, -order, -class2)
# 
# m_com5 <- as.matrix(com5) 
# 
# set.seed(123)
# nmds <- metaMDS(m_com5, distance = "euclidean")
# nmds
# 
# # extract NMDS scores (x and y coordinates)
# filled_scores5 <- as.data.frame(scores(nmds)$sites)
# 
# # add columns to data frame
# filled_scores5$class2 <- com_data5$class2
# filled_scores5$species <- com_data5$species
# filled_scores5$order <- com_data5$order
# 
# filled_scores5 <- filled_scores5 %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))
# 
# species_scores <- as.data.frame(nmds[["species"]]) %>%
#   rownames_to_column(var = "var")
# 
# 
# nmds_plot5 <- ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   # theme(legend.position = "none") +
#   labs(x = "NMDS1", y = "NMDS2") +
#   stat_ellipse(geom = "polygon", aes(fill = class2, color = class2), alpha = 0.2) +
#   geom_segment(data = species_scores, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd = 0.3) +
#   ggrepel::geom_text_repel(data = species_scores, aes(x = MDS1, y = MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
# 
# nmds_plot5
# 
# #ggsave(nmds_plot5, file = paste0("nmds_19.png"), path = here::here("04_figs", "nmds"), height = 12, width = 12)
# 
# dst5 <- dist(com5, method = "euclidean")
# 
# bd5 <- betadisper(dst5, com_data5$class2)
# 
# anova(bd5)
# 
# permutest(bd5, pairwise = TRUE, permutations = 9999)

```


```{r}
# continuous

# plot1_data = fishbase_data

# #option1 sort by variable
# plot1_data = plot1_data %>%
#   dplyr::select(r_fin, species, class2) %>%
#   mutate(r_fin = sort(r_fin)) %>%
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(species = fct_relevel(plot1_data$species))
#
# ggplot(plot1_data, aes(species, r_fin)) +
#   geom_point(aes(color = class2))


# sort by variable and color by class

# plot2_data = filled_data2 %>%
#   filter(temp_preferred != "NA")
#
# plot2_data = filled_data2
#
# plot2_data <- plot2_data %>%
#   #mutate(length = sort(length)) %>%
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(order = as.factor(order)) %>%
#   mutate(species = fct_relevel(plot2_data$species))
#
# #more descriptive**** # USE THESE PLOTS for continuous
# ind3 <- ggplot(plot2_data, aes(r_fin, class2)) +
#   geom_boxplot(aes(color = class2)) +
#   theme_bw(base_size  = 20) +
#   labs(y = "Movement Class",
#        x = "Growth Rate", color = "Class (homerange/PLD)")
#   #theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
#   #theme(legend.position = "none")
# ind3
#
#
# ggsave(ind3, file = paste0("ind3.png"), path = here::here("04_figs", "individual_plots"))


# #more to find groupings
# ggplot(plot1_data, aes(species, r_fin)) +
#   geom_point(aes(color = class2))
```


```{r}
# plot2_data = filled_data2
#
# plot2_data <- plot2_data %>%
#   #mutate(length = sort(length)) %>%
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(movement_keyword = as.factor(movement_keyword)) %>%
#   mutate(species = fct_relevel(plot2_data$species))
#
# #more descriptive**** # USE THESE PLOTS for continuous
# ind2 <- ggplot(plot2_data, aes(movement_keyword, class2)) +
#   geom_jitter(aes(color = class2)) +
#   theme_bw(base_size  = 20) +
#   labs(y = "Movement Class",
#        x = "Movement Keyword", color = "Class (homerange/PLD)") +
#   theme(axis.text.x = element_text(angle = 30, hjust=1, size = 8))
#   #theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
#   #theme(legend.position = "none")
# ind2
#
#
# ggsave(ind2, file = paste0("ind2.png"), path = here::here("04_figs", "individual_plots"))
```


```{r}
# categorical
# plot2_data = filled_data %>%
#   dplyr::select(demers_pelag, r_fin, species, class2, family, movement_keyword,
#          order, class_taxon, longevity_wild, diet_troph3) %>%
#   mutate(class2 = as.factor(class2), order = as.factor(order))
#
# # #to find groupings
# # ggplot(plot2_data, aes(class2, demers_pelag)) +
# #   geom_point(aes(color = family)) +
# #   theme(legend.position = "none")
#
# #more descriptive*** - USE THIS FOR CATEGORICAL
# ggplot(plot2_data, aes(class2, movement_keyword)) +
#   #geom_point(aes(color = demers_pelag)) +
#   geom_boxplot(aes(color = class2)) +
#   theme_bw() +
#   labs(x = "Movement Keyword",
#        y = "Movement Class") +
#   #theme(legend.position = "none") +
#   scale_color_viridis_d() +
#   theme(axis.text.x = element_text(angle = 30, hjust=1))
#   #scale_color_manual(values = c("", ""))
```


### Metabolic data

```{r}
# plot2_data = fishbase_data
# plot2_data <- plot2_data[!is.na(plot2_data$oxygen_cons),]
#
# ggplot(plot2_data, aes(oxygen_cons, class2)) +
#   geom_point(aes(color = class2))
#
# #option1 sort by oxygen_cons
# plot2_data = plot2_data %>%
#   dplyr::select(oxygen_cons, species, class2, genus, order, temp_preferred) %>%
#   mutate(oxygen_cons = sort(oxygen_cons)) %>%
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(species = fct_relevel(plot2_data$species))
#
# ggplot(plot2_data, aes(oxygen_cons, class2)) +
#   geom_jitter(aes(color = oxygen_cons)) +
#   theme_bw() +
#   theme(legend.position="none") +
#   theme(axis.text.x=element_blank()) +
#   labs(x = "Species",
#        y = "O2 Consumption Rate") # +
#   #scale_color_viridis_d()
#
# # no noteworthy groupings for class - try coloring by family/order
#
#
# #option 2 sort by class
# plot2_data = plot2_data %>%
#   dplyr::select(oxygen_cons, species, class2, order, temp_preferred)%>%
#   mutate(class2 = sort(class2)) %>%
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(species = fct_relevel(plot2_data$species))
#
# #more descriptive****
# ggplot(plot2_data, aes(temp_preferred, class2)) +
#   geom_point(aes(color = class2)) +
#   theme_bw() +
#   labs(x = "Optimal Temp",
#        y = "Movement Class") +
#   theme(legend.position = "none")
#
```




```{r}
#
# plot3_data = fishbase_data %>%
#   dplyr::select(demers_pelag, env_temp, body_shape_i, species, class, class2) %>%
#   mutate(class = as.factor(class), class2 = as.factor(class2))
#
# #to find groupings
# ggplot(plot3_data, aes(species, env_temp)) +
#   geom_point(aes(color = class2))
#
# #more descriptive
# ggplot(plot3_data, aes(body_shape_i, class2)) +
#   geom_jitter(aes(color = body_shape_i)) +
#   theme_bw() +
#   labs(x = "env_temp",
#        y = "Movement Class",
#        color = "Habitat Type") +
#   scale_color_viridis_d() +
#   theme(axis.text.x = element_text(angle = 30, hjust=1))
#   #scale_color_manual(values = c("", ""))
#
# plot4_data = fishbase_data %>%
#   dplyr::select(length, species, class2) %>%
#   #mutate(class2 = sort(length)) %>% this line is causing problemssssss
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(species = fct_relevel(plot4_data$species))
#
# ggplot(plot4_data, aes(species, length)) +
#   geom_jitter(aes(color = class2))
#
```


### DO NMDS HERE INSTEAD OF PCA WITH EUCLIDEAN DISTANCE FROM NOW ON 


Most of this can probably go in old now and then doing overall NMDS and FAMD then subset ones

### PCA with new variables added

```{r}
# fishbase_p1 <- fishbase_data %>%
#   dplyr::select(where(is.numeric)) %>%
#   dplyr::select(r_fin, longevity_wild, length, fecundity_mean, diet_troph3, class2) %>%
#   na.omit() # YOU WANT TO REMOVE COLUMNS WITH NAS FOR PCA, AND PROBABLY NMDS TOO
#
# p1 <- PCA(fishbase_p1[,1:5], graph = FALSE) # runs the PCA on the first 5 columns (not class)
# p1
#
# eigen_p1 <- p1$eig
# eigen_p1 # 69.07% variance explained by first 2 components
#
# p1_fig <- fviz_pca_ind(p1, label = "none", habillage = fishbase_p1$class2, addEllipses = TRUE,
#                 ellipse.level = 0.95) +
#             theme_minimal()
# p1_fig
#
# ggsave(p1_fig, file = paste0("p1_fig.png"), path = here::here("04_figs", "current_pca_famd"))
#
#
# # Vizualize variable contributions to components
#
# fviz_pca_biplot(p1, repel = TRUE, label = "none", habillage = fishbase_p1$class2)
#
#
# # Contributions of variables to PC1
# p1_c1 <- fviz_contrib(p1, choice = "var", axes = 1, top = 10)
#
# # Contributions of variables to PC2
# p1_c2 <- fviz_contrib(p1, choice = "var", axes = 2, top = 10)
#
# components_p1 <- gridExtra::grid.arrange(p1_c1, p1_c2, ncol = 2)
#
#
# #ggsave(components_p1, file = paste0("components_p1.png"), path = here::here("04_figs", "current_pca_famd"))
```



```{r}
# fishbase_p2 <- fishbase_data %>%
#   dplyr::select(where(is.numeric)) %>%
#   dplyr::select(r_fin, length, diet_troph3, class2) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   na.omit()
#
# p2 <- PCA(fishbase_p2[,1:3], graph = FALSE) # runs the PCA on the first 4 columns (not class)
# p2
#
# eigen_p2 <- p2$eig
# eigen_p2 # 74.71% of variance explained by first 2 components
#
# p2_fig <- fviz_pca_ind(p2, label = "none", habillage = fishbase_p2$class2, addEllipses = TRUE,
#                 ellipse.level = 0.95) +
#             theme_minimal()
# p2_fig

# ggsave(p2_fig, file = paste0("p2_fig.png"), path = here::here("04_figs"))
```



### FAMD with new variables added

```{r}
# fishbase_f1 <- fishbase_data %>%
#    dplyr::select(weight, env_temp, temp_preferred, fecundity_mean, r_fin, class2) %>%
#    mutate(class2 = as.factor(class2),
#           env_temp = as.factor(env_temp)) %>%
#    na.omit()
#
# f1 <- FAMD(fishbase_f1[,1:5], graph=FALSE)
# f1
#
# eigen_f1 <- f1$eig
# eigen_f1 # 42.11% of variance explained by first 2 components
#
#
# f1_fig <- fviz_famd_ind(f1, label = "none", habillage = fishbase_f1$class2, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# f1_fig
#
# # ggsave(f1_fig, file = paste0("f1_fig.png"), path = here::here("04_figs"))
```

```{r}
#
# fishbase_f2 <- fishbase_data %>%
#    dplyr::select(r_fin, env_temp, temp_preferred, demers_pelag,
#           body_shape_i, fecundity_mean, longevity_wild, movement_keyword, length, class2, kfin, diet_troph3, food_i) %>%
#    mutate(class2 = as.factor(class2),
#           env_temp = as.factor(env_temp),
#           demers_pelag = as.factor(demers_pelag),
#           body_shape_i = as.factor(body_shape_i)) %>%
#    na.omit()
#
# f2 <- FAMD(fishbase_f2[,1:12], graph=FALSE)
# f2
#
# eigen_f2 <- f2$eig
# eigen_f2 # 42.11% of variance explained by first 2 components
#
#
# f2_fig <- fviz_famd_ind(f2, label = "none", habillage = fishbase_f2$class2, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# f2_fig
#
# fviz_famd_var(f2, geom = "arrow", )


# ggsave(f2_fig, file = paste0("f1_fig.png"), path = here::here("04_figs"))
```


```{r}
# fishbase_f3 <- fishbase_data %>%
#   dplyr::select(length, diet_troph3, demers_pelag, order) %>%
#   mutate(order = as.factor(order), demers_pelag = as.factor(demers_pelag)) %>%
#   na.omit()
#
# f3 <- FAMD(fishbase_f3[,1:3], graph=FALSE)
# f3
#
# eigen_f3 <- f3$eig
# eigen_f3 # 30.38% of variance explained by first 2 components
#
#
# f3_fig <- fviz_famd_ind(f3, label = "none", habillage = fishbase_f3$order, addEllipses = FALSE, ellipse.level = 0.95) +
#   theme_minimal() +
#   theme(legend.position = "none")
# f3_fig


# ggsave(f3_fig, file = paste0("f3_fig.png"), path = here::here("04_figs", "current_pca_famd"))
```


### Initial PCA and FAMD

```{r}
# fishbase_pca1 <- fishbase_data %>%
#   dplyr::select(where(is.numeric)) %>%
#   dplyr::select(r_fin, kfin, length, diet_troph3, class) %>%
#   mutate(class = as.factor(class))
#
# # this will only work if all variables are numeric
# #cor.mat <- round(cor(fishbase_pca1),2) # nothing is highly correlated - good!
# #cor.mat
#
# pca1 <- PCA(fishbase_pca1[,1:4], graph = FALSE) # runs the PCA on the first 4 columns (not class)
# pca1
#
# eigen_pca1 <- pca1$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
# eigen_pca1 # 66.4% of variance explained by first 2 components
#
# fviz_screeplot(pca1, ncp=10) # scree plot shows eigenvalues/variances associated w components
#
#
# plot(pca1, choix = "var") # this is the same plot as the original PCA code on line 31?
#
#
# pca1_fig <- fviz_pca_ind(pca1, label = "none", habillage = fishbase_pca1$class, addEllipses = TRUE,
#                 ellipse.level = 0.95) +
#             theme_minimal()
# pca1_fig

# ggsave(pca1_fig, file = paste0("pca1_fig.png"), path = here::here("04_figs"))
```


```{r}
#
# fishbase_pca2 <- fishbase_data %>%
#   dplyr::select(where(is.numeric)) %>%
#   dplyr::select(r_fin, length, diet_troph3, class) %>%
#   mutate(class = as.factor(class))
#
# pca2 <- PCA(fishbase_pca2[,1:3], graph = FALSE) # runs the PCA on the first 4 columns (not class)
#
# eigen_pca2 <- pca2$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
# eigen_pca2 # 81.45% of variance explained by first 2 components
#
# plot(pca2, choix = "var")
#
#
# pca2_fig <- fviz_pca_ind(pca2, label = "none", habillage = fishbase_pca2$class, addEllipses = TRUE,
#                 ellipse.level = 0.95) +
#             theme_minimal()
# pca2_fig

# ggsave(pca2_fig, file = paste0("pca2_fig.png"), path = here::here("04_figs"))
```


```{r}
#
# fishbase_pca3 <- fishbase_data %>%
#   dplyr::select(where(is.numeric)) %>%
#   dplyr::select(kfin, length, diet_troph3, class) %>%
#   mutate(class = as.factor(class))
#
# pca3 <- PCA(fishbase_pca3[,1:3], graph = FALSE) # runs the PCA on the first 4 columns (not class)
#
# eigen_pca3 <- pca3$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
# eigen_pca3 # 81.29% of variance explained by first 2 components
#
# plot(pca3, choix = "var")
#
#
# pca3_fig <- fviz_pca_ind(pca3, label = "none", habillage = fishbase_pca3$class, addEllipses = TRUE,
#                 ellipse.level = 0.95) +
#             theme_minimal()
# pca3_fig
#
# #ggsave(pca3_fig, file = paste0("pca3_fig.png"), path = here::here("04_figs"))
```


```{r}
# fishbase_pca4 <- fishbase_data %>%
#   dplyr::select(where(is.numeric)) %>%
#   dplyr::select(length, diet_troph3, class) %>%
#   mutate(class = as.factor(class))
#
# pca4 <- PCA(fishbase_pca4[,1:2], graph = FALSE) # runs the PCA on the first 4 columns (not class)
#
# eigen_pca4 <- pca4$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
# eigen_pca4 # only 2 components
#
# plot(pca4, choix = "var")
#
#
# pca4_fig <- fviz_pca_ind(pca4, label = "none", habillage = fishbase_pca4$class, addEllipses = TRUE,
#                 ellipse.level = 0.95) +
#             theme_minimal()
# pca4_fig # group 6 pretty clearly defined?
#
# #ggsave(pca4_fig, file = paste0("pca4_fig.png"), path = here::here("04_figs"))
```


### FAMD

```{r}
#
# fishbase_famd1 <- fishbase_data %>%
#   dplyr::select(r_fin, kfin, length, diet_troph3, movement_keyword, food_i, class) %>%
#   mutate(class = as.factor(class), movement_keyword = as.factor(movement_keyword), food_i = as.factor(food_i)) %>%
#   na.omit(fishbase_famd1)
#
#
# famd1 <- FAMD(fishbase_famd1[,1:6], graph=FALSE)
# famd1
#
# eigen_famd1 <- famd1$eig
# eigen_famd1 # 7.23% of variance explained by first 2 components!?
#
#
# famd1_fig <- fviz_famd_ind(famd1, label = "none", habillage = fishbase_famd1$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# famd1_fig
# #ggsave(famd1_fig, file = paste0("famd1_fig.png"), path = here::here("04_figs"))
```


```{r}
#
# fishbase_famd2 <- fishbase_data %>%
#   dplyr::select(length, food_i, class) %>%
#   mutate(class = as.factor(class), food_i = as.factor(food_i)) %>%
#   na.omit(fishbase_famd2)
#
# famd2 <- FAMD(fishbase_famd2[,1:2], graph=FALSE)
#
# eigen_famd2 <- famd2$eig
# eigen_famd2 # 5.49%
#
#
# famd2_fig <- fviz_famd_ind(famd2, label = "none", habillage = fishbase_famd2$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# famd2_fig
# #ggsave(famd2_fig, file = paste0("famd2_fig.png"), path = here::here("04_figs"))
```


```{r}
#
# fishbase_famd3 <- fishbase_data %>%
#   dplyr::select(r_fin, kfin, diet_troph3, iucn_category, class) %>%
#   mutate(class = as.factor(class), iucn_category = as.factor(iucn_category)) %>%
#   na.omit(fishbase_famd3)
#
# famd3 <- FAMD(fishbase_famd3[,1:4], graph=FALSE)
# famd3
#
# eigen_famd3 <- famd3$eig
# eigen_famd3 # 31.50% of variance explained by first 2 components
#
#
# famd3_fig <- fviz_famd_ind(famd3, label = "none", habillage = fishbase_famd3$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# famd3_fig
# #ggsave(famd3_fig, file = paste0("famd3_fig.png"), path = here::here("04_figs"))
```


```{r}
#
# fishbase_famd4 <- fishbase_data %>%
#   dplyr::select(length, iucn_category, class) %>%
#   mutate(class = as.factor(class), iucn_category = as.factor(iucn_category)) %>%
#   na.omit(fishbase_famd4)
#
# famd4 <- FAMD(fishbase_famd4[,1:2], graph=FALSE)
# famd4
#
# eigen_famd4 <- famd4$eig
# eigen_famd4 # 38.75% of variance explained by first 2 components
#
#
# famd4_fig <- fviz_famd_ind(famd4, label = "none", habillage = fishbase_famd4$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# famd4_fig
# #ggsave(famd4_fig, file = paste0("famd4_fig.png"), path = here::here("04_figs"))
```


```{r}
# fishbase_famd5 <- fishbase_data %>%
#   dplyr::select(length, demers_pelag, diet_troph3, class) %>%
#   mutate(class = as.factor(class),
#          demers_pelag = as.factor(demers_pelag),
#          diet_troph3 = as.factor(diet_troph3)) %>%
#   na.omit(fishbase_famd5)
#
# famd5 <- FAMD(fishbase_famd5[,1:3], graph=FALSE)
# famd5
#
# eigen_famd5 <- famd5$eig
# eigen_famd5 # 1.27% of variance explained by first 2 components!?
#
#
# famd5_fig <- fviz_famd_ind(famd5, label = "none", habillage = fishbase_famd5$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
# famd5_fig
#
# ggsave(famd5_fig, file = paste0("famd5_fig.png"), path = here::here("04_figs"))
```


```{r}
# fishbase_famd6 <- fishbase_data %>%
#   dplyr::select(length, diet_troph3, demers_pelag, class) %>%
#   mutate(class = as.factor(class), demers_pelag = as.factor(demers_pelag)) %>%
#   na.omit()
#
# famd6 <- FAMD(fishbase_famd6[,1:3], graph=FALSE)
# famd6
#
# eigen_famd6 <- famd6$eig
# eigen_famd6 # 30.38% of variance explained by first 2 components
#
#
# famd6_fig <- fviz_famd_ind(famd6, label = "none", habillage = fishbase_famd6$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
# famd6_fig # still pretty good groupings here
#
# famd6_1_fig <- fviz_famd_ind(famd6, label = "none", habillage = fishbase_famd6$class) +
#   theme_minimal()
# famd6_1_fig # really interesting.. what does it mean?
#
#
# #ggsave(famd6_fig, file = paste0("famd6_fig.png"), path = here::here("04_figs"))
# #ggsave(famd6_1_fig, file = paste0("famd6_1_fig.png"), path = here::here("04_figs"))
```



```{r}
#
# fishbase_famd5 <- fishbase_data %>%
#   dplyr::select(r_fin, kfin, length, diet_troph3, movement_keyword, demers_pelag, food_i, class) %>%
#   mutate(class = as.factor(class),
#          movement_keyword = as.factor(movement_keyword),
#          food_i = as.factor(food_i),
#          demers_pelag = as.factor(demers_pelag)) %>%  # error here?  <--------------------------------------------
#   na.omit(fishbase_famd5)
#
#
# famd5 <- FAMD(fishbase_famd5[,1:7], graph=FALSE)
# famd5
#
# eigen_famd5 <- famd5$eig
# eigen_famd5 # 7.23% of variance explained by first 2 components!?
#
#
# famd5_fig <- fviz_famd_ind(famd5, label = "none", habillage = fishbase_famd5$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
#
# famd5_fig
# #ggsave(famd5_fig, file = paste0("famd5_fig.png"), path = here::here("04_figs"))
#
```


```{r}
#
# fishbase_famd5 <- fishbase_data %>%
#   dplyr::select(r_fin, kfin, length, diet_troph3, food_i, movement_keyword, iucn_category, demers_pelag, class) %>%
#   mutate(class = as.factor(class),
#          demers_pelag = as.factor(demers_pelag),
#          iucn_category = as.factor(iucn_category),
#          food_i = as.factor(food_i),
#          movement_keyword = as.factor(movement_keyword)) %>%
#   na.omit(fishbase_famd5) # error why?  <------------------------------------------------------------------
#
# famd5 <- FAMD(fishbase_famd5[,1:8], graph=FALSE)
# famd5
#
# eigen_famd5 <- famd5$eig
# eigen_famd5 # 30.38% of variance explained by first 2 components
#
#
# famd5_fig <- fviz_famd_ind(famd5, label = "none", habillage = fishbase_famd5$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
# famd5_fig
#
#
# #ggsave(famd5_fig, file = paste0("famd5_fig.png"), path = here::here("04_figs"))
#
```


Sort by variable and color by class

```{r}
#
# plot2_data = filled_data %>%
#   filter(r_fin != "NA")
#
# #plot2_data = filled_data
#
# plot2_data <- plot2_data %>%
#   #mutate(r_fin = sort(r_fin)) %>%
#   mutate(species = as.factor(species)) %>%
#   mutate(class2 = as.factor(class2)) %>%
#   mutate(species = fct_relevel(plot2_data$species))
#
# ggplot(plot2_data, aes(species, r_fin)) +
#   geom_point(aes(color = class2)) +
#   theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
#   theme(legend.position = "none")
#
```
