---
title: "PCA"
output: html_document
date: "2023-04-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(rfishbase)
library(FactoMineR)
library(factoextra)
library(vcd)
library(purrr)
library(car)
library(vegan)
```

### Load data

```{r}
fishbase_data <- read_csv(here::here("01_data", "02_processed_data", "fishbase_data.csv")) %>% 
  mutate_if(is.character, as.factor)
filled_data <- read_csv(here::here("01_data", "02_processed_data", "filled_data.csv")) %>% 
  mutate_if(is.character, as.factor)
filled_data2 <- read_csv(here::here("01_data", "02_processed_data", "filled_data2.csv")) %>% 
  mutate_if(is.character, as.factor)
```

### Plot data


```{r}

# continuous

plot_data = fishbase_data %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order))

response = "class2"
expl_cont = c("temp_preferred", "fecundity_max", "fecundity_mean",
         "longevity_wild", "sp_richness", "length", "kfin", "r_fin", "diet_troph3")

scatter_data = function(x, y) {
     ggplot(plot_data, aes(x = .data[[x]], y = .data[[y]]) ) +
          geom_jitter(aes(color = order)) +
          theme_bw() +
          theme(legend.position = "none")
}

#scatter_data(x = "temp_preferred", y = "class2")

plots_cont = map(expl_cont, ~scatter_data(.x, "class2"))
plots_cont

```


```{r}

# categorical

expl_cat = c("climate", "body_shape_i", "dev_place", "movement_keyword", 
             "demers_pelag", "iucn_category", "food_i")

plots_cat = map(expl_cat, ~scatter2_data(.x, "class2"))
plots_cat


```


### Look at filled data

```{r}

# continuous

plot_filled = filled_data %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order))

expl_cont_filled = c("temp_preferred", "fecundity_max", "fecundity_mean",
         "longevity_wild", "sp_richness", "length", "kfin", "r_fin", "diet_troph3")

scatter_filled = function(x, y) {
     ggplot(plot_filled, aes(x = .data[[x]], y = .data[[y]]) ) +
          geom_jitter(aes(color = class2)) +
          theme_bw() +
          theme(legend.position = "none")
}

plots_cont_filled = map(expl_cont_filled, ~scatter_filled(.x, "class2"))
plots_cont_filled

scatter_filled(x = "length", y = "class2")


```


```{r}

# categorical

expl_cat_filled = c("climate", "body_shape_i", "dev_place", "movement_keyword", "demers_pelag",
                    "iucn_category", "food_i")

plots_cat_filled = map(expl_cat_filled, ~scatter_filled(.x, "class2"))
plots_cat_filled


```


### Look at filled data including oxygen_cons and temp_range

```{r}

# continuous

plot_filled2 = filled_data2 %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order))

expl_cont_filled2 = c("temp_preferred", "fecundity_max", "fecundity_mean",
         "longevity_wild", "sp_richness", "length", "kfin", "r_fin", "diet_troph3",
         "oxygen_cons", "temp_range")

scatter_filled2 = function(x, y) {
     ggplot(plot_filled2, aes(x = .data[[x]], y = .data[[y]]) ) +
          geom_jitter(aes(color = class2)) +
          theme_bw() +
          theme(legend.position = "none")
}

plots_cont_filled2 = map(expl_cont_filled2, ~scatter_filled2(.x, "class2"))
plots_cont_filled2



ind1_1 <- scatter_filled2(x = "length", y = "class2")
ind1_1


```

Sort by variable and color by class

```{r}

plot2_data = filled_data2 %>% 
  filter(r_fin != "NA") 

#plot2_data = filled_data2

plot2_data <- plot2_data %>%  
  #mutate(r_fin = sort(r_fin)) %>% 
  mutate(species = as.factor(species)) %>% 
  mutate(class2 = as.factor(class2)) %>% 
  mutate(species = fct_relevel(plot2_data$species))

ggplot(plot2_data, aes(species, r_fin)) +
  geom_point(aes(color = class2)) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  theme(legend.position = "none")

```


### Check Variable Inflation Factors (VIF)

none are above 5

```{r}

model <- lm(class2 ~ temp_preferred + fecundity_mean + #fecundity_max + longevity_wild + 
              sp_richness + length + kfin + r_fin + diet_troph3, 
            data = filled_data)

summary(model)

vif(model) # when including both fecundity variables, they have a VIF over 5

#create vector of VIF values
vif_values <- vif(model)

#create horizontal bar chart to display each VIF value
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue", xlim = c(0,6), las = 2)

#add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)


#model2 <- lm(class2 ~ temp_preferred + length + r_fin, data = filled_data)
#summary(model)

```


### Run nMDS and FAMD


#### FAMD

```{r}

# getting error when I include demers_pelag! non-unique row value when setting row names?


famd_filled <- filled_data %>% 
   select(temp_preferred, fecundity_mean, #fecundity_max, longevity_wild, 
          sp_richness, 
          length, kfin, r_fin, diet_troph3, climate, body_shape_i, dev_place, 
          movement_keyword, #demers_pelag, iucn_category, 
          food_i, class2) %>% 
   mutate(class2 = as.factor(class2)) %>% 
   na.omit()
 

f1 <- FAMD(famd_filled[,1:12], graph=FALSE)
f1

eigen_f1 <- f1$eig
eigen_f1 # only 8% of variance explained by first 2 components......


f1_fig <- fviz_famd_ind(f1, label = "none", habillage = famd_filled$class2, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

f1_fig

#ggsave(f1_fig, file = paste0("f1.png"), path = here::here("04_figs", "current_pca_famd"))

```


```{r}

famd_filled2 <- filled_data %>% 
   select(fecundity_mean, length, r_fin, diet_troph3, 
          #climate, 
          body_shape_i, 
          dev_place, 
          movement_keyword, 
          #demers_pelag, 
          #iucn_category, 
          #food_i, 
          class2) %>% 
   mutate(class2 = as.factor(class2)) %>% 
   na.omit()
 

f2 <- FAMD(famd_filled2[,1:8], graph=FALSE)
f2

eigen_f2 <- f2$eig
eigen_f2 


f2_fig <- fviz_famd_ind(f2, label = "none", habillage = famd_filled2$class2, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_bw()

f2_fig # 2 and 3 completely separate, 2 and 5 as well

ggsave(f2_fig, file = paste0("f2.png"), path = here::here("04_figs", "current_pca_famd"))

```


```{r}

famd_filled3 <- filled_data2 %>% 
   select(fecundity_mean, length, r_fin, diet_troph3,
          oxygen_cons,
          #temp_range,
          climate, 
          #body_shape_i, 
          #dev_place, 
          movement_keyword, 
          #demers_pelag, 
          #iucn_category, 
          #food_i, 
          class2) %>% 
   mutate(class2 = as.factor(class2)) %>% 
   na.omit()
 

f3 <- FAMD(famd_filled3[,1:7], graph=FALSE)
f3

eigen_f3 <- f3$eig
eigen_f3 


f3_fig <- fviz_famd_ind(f3, label = "none", habillage = famd_filled3$class2, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_bw()

f3_fig # 2 and 3 completely separate, 2 and 5 as well

#ggsave(f3_fig, file = paste0("f3.png"), path = here::here("04_figs", "current_pca_famd"))

```


```{r}

famd_filled4 <- filled_data2 %>% 
   select(fecundity_mean, 
          length, 
          r_fin, 
          #diet_troph3,
          oxygen_cons,
          temp_range,
          #climate, 
          #body_shape_i, 
          #dev_place, 
          movement_keyword, 
          #demers_pelag, 
          #iucn_category, 
          #food_i, 
          class2) %>% 
   mutate(class2 = as.factor(class2)) %>% 
   na.omit()
 

f4 <- FAMD(famd_filled4[,1:6], graph=FALSE)
f4

eigen_f4 <- f4$eig
eigen_f4 


f4_fig <- fviz_famd_ind(f4, label = "none", habillage = famd_filled4$class2, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

f4_fig # 2 and 3 completely separate, 2 and 5 as well

#ggsave(f4_fig, file = paste0("f4.png"), path = here::here("04_figs", "current_pca_famd"))

```



#### nMDS

```{r}

com_data <- filled_data %>% 
   select(temp_preferred, fecundity_mean, #fecundity_max, longevity_wild, 
          sp_richness, length, kfin, r_fin, diet_troph3, 
          species, order, class2) %>% 
   filter(temp_preferred > 0) %>% 
   na.omit()

com <- com_data %>% 
  select(-species, -order, -class2)

m_com = as.matrix(com) # make matrix from dataframe

set.seed(123)
nmds = metaMDS(m_com, distance = "euclidean")
nmds # stress: 0.1716711, weak ties

#extract NMDS scores (x and y coordinates)
filled_scores = as.data.frame(scores(nmds)$sites)

#add columns to data frame 
filled_scores$class2 = com_data$class2
filled_scores$species = com_data$species
filled_scores$order = com_data$order

filled_scores <- filled_scores %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))

species_scores = as.data.frame(nmds[["species"]]) %>% 
  rownames_to_column(var = "var")

nmds_plot1 = ggplot(filled_scores, aes(x = NMDS1, y = NMDS2)) + 
  geom_point(aes(color = class2)) +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2), alpha = 0.2) +
  geom_segment(data = species_scores, aes(x = 0, xend=MDS1, y=0, yend=MDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) +
  ggrepel::geom_text_repel(data = species_scores, aes(x=MDS1, y=MDS2, label = var), cex = 3, direction = "both", segment.size = 0.25)
  

nmds_plot1 # classes 3 and 4 are distinct!
# coloring by order gave four plus distinct groups..?

#ggsave(nmds_plot1, file = paste0("nmds_plot1.png"), path = here::here("04_figs"))


```

##### another

```{r}

com_data2 <- filled_data %>% 
   select(temp_preferred, #fecundity_mean, #fecundity_max, 
          longevity_wild, 
          #sp_richness, 
          length, kfin, r_fin, diet_troph3, 
          species, order, class2) %>% 
   filter(temp_preferred > 0) %>% 
   na.omit()

com2 <- com_data2 %>% 
  select(-species, -order, -class2)

m_com2 = as.matrix(com2) # make matrix from dataframe

set.seed(123)
nmds = metaMDS(m_com2, distance = "euclidean")
nmds # stress: 0.1432603, weak ties


#extract NMDS scores (x and y coordinates)
filled_scores2 = as.data.frame(scores(nmds)$sites)

#add columns to data frame 
filled_scores2$class2 = com_data2$class2
filled_scores2$species = com_data2$species
filled_scores2$order = com_data2$order

filled_scores2 <- filled_scores2 %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))


nmds_plot2 = ggplot(filled_scores2, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  #theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2), alpha = 0.2)+
  geom_segment(data = filled_scores2, aes(x = 0, xend=NMDS1, y=0, yend=NMDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) +
  ggrepel::geom_text_repel(data = filled_scores2, aes(x=NMDS1, y=NMDS2, label = class2), cex = 3, direction = "both", segment.size = 0.25)

nmds_plot2 # 2 and 3 are pretty distinct, 3 and 7 as well

#ggsave(nmds_plot2, file = paste0("nmds_plot2.png"), path = here::here("04_figs"))

```

##### another

```{r}

com_data3 <- filled_data %>% 
   select(temp_preferred, #fecundity_mean, 
          fecundity_max, 
          #longevity_wild, 
          #sp_richness, 
          length, kfin, r_fin, diet_troph3, 
          species, order, class2) %>% 
   filter(temp_preferred > 0) %>% 
   na.omit()

com3 <- com_data3 %>% 
  select(-species, -order, -class2)

m_com3 = as.matrix(com3) # make matrix from dataframe

set.seed(123)
nmds = metaMDS(m_com3, distance = "euclidean")
nmds # stress: 0.1607562, weak ties


#extract NMDS scores (x and y coordinates)
filled_scores3 = as.data.frame(scores(nmds)$sites)

#add columns to data frame 
filled_scores3$class2 = com_data3$class2
filled_scores3$species = com_data3$species
filled_scores3$order = com_data3$order

filled_scores3 <- filled_scores3 %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))


nmds_plot3 = ggplot(filled_scores3, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  #theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2), alpha = 0.2)

nmds_plot3 # 2 and 3 are pretty distinct, 3 and 4 as well

#ggsave(nmds_plot3, file = paste0("nmds_plot3.png"), path = here::here("04_figs"))

```

##### another

```{r}

com_data4 <- filled_data %>% 
   select(temp_preferred, #fecundity_mean, 
          #fecundity_max, 
          #longevity_wild, 
          #sp_richness, 
          length, kfin, r_fin, diet_troph3, 
          species, order, class2) %>% 
   filter(temp_preferred > 0) %>% 
   na.omit()

com4 <- com_data4 %>% 
  select(-species, -order, -class2)

m_com4 = as.matrix(com4) # make matrix from dataframe

set.seed(123)
nmds = metaMDS(m_com4, distance = "euclidean")
nmds # stress: 0.1343541, weak ties


#extract NMDS scores (x and y coordinates)
filled_scores4 = as.data.frame(scores(nmds)$sites)

#add columns to data frame 
filled_scores4$class2 = com_data4$class2
filled_scores4$species = com_data4$species
filled_scores4$order = com_data4$order

filled_scores4 <- filled_scores4 %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))


nmds_plot4 = ggplot(filled_scores4, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  #theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2), alpha = 0.2)

nmds_plot4 # 2, 3, 4, 7 

#ggsave(nmds_plot4, file = paste0("nmds_plot4.png"), path = here::here("04_figs"))

```

##### another

```{r}

com_data5 <- filled_data %>% 
   select(temp_preferred, 
          fecundity_mean, 
          #fecundity_max, 
          #longevity_wild, 
          #sp_richness, 
          #length, 
          #kfin, 
          r_fin, 
          diet_troph3, 
          species, order, class2) %>% 
   filter(temp_preferred > 0) %>% 
   na.omit()

com5 <- com_data5 %>% 
  select(-species, -order, -class2)

m_com5 = as.matrix(com5) # make matrix from dataframe

set.seed(123)
nmds = metaMDS(m_com5, distance = "euclidean")
nmds # stress: 0.08614346, weak ties


#extract NMDS scores (x and y coordinates)
filled_scores5 = as.data.frame(scores(nmds)$sites)

#add columns to data frame 
filled_scores5$class2 = com_data5$class2
filled_scores5$species = com_data5$species
filled_scores5$order = com_data5$order

filled_scores5 <- filled_scores5 %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))


nmds_plot5 = ggplot(filled_scores5, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  #theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2), alpha = 0.2)

nmds_plot5 # 2, 3, 4

#ggsave(nmds_plot5, file = paste0("nmds_plot5.png"), path = here::here("04_figs"))

```


##### another

```{r}

com_data6 <- filled_data %>% 
   select(#temp_preferred, 
          fecundity_mean, 
          #fecundity_max, 
          #longevity_wild, 
          #sp_richness, 
          length, 
          #kfin, 
          r_fin, 
          diet_troph3, 
          species, order, class2) %>% 
   #filter(temp_preferred > 0) %>% 
   na.omit()

com6 <- com_data6 %>% 
  select(-species, -order, -class2)

m_com6 = as.matrix(com6) # make matrix from dataframe

set.seed(123)
nmds = metaMDS(m_com6, distance = "euclidean")
nmds # stress: 0.05406095, weak ties


#extract NMDS scores (x and y coordinates)
filled_scores6 = as.data.frame(scores(nmds)$sites)

#add columns to data frame 
filled_scores6$class2 = com_data6$class2
filled_scores6$species = com_data6$species
filled_scores6$order = com_data6$order

filled_scores6 <- filled_scores6 %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order), species = as.factor(species))


nmds_plot6 = ggplot(filled_scores6, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = class2)) +
  #theme(legend.position = "none") +
  labs(x = "NMDS1", y = "NMDS2") +
  stat_ellipse(geom = "polygon", aes(fill = class2), alpha = 0.2)

nmds_plot6 # 2, 3, 4

#ggsave(nmds_plot6, file = paste0("nmds_plot6.png"), path = here::here("04_figs"))

```










```{r}
#continuous 

# plot1_data = fishbase_data 

# #option1 sort by variable
# plot1_data = plot1_data %>% 
#   select(r_fin, species, class2) %>% 
#   mutate(r_fin = sort(r_fin)) %>% 
#   mutate(species = as.factor(species)) %>% 
#   mutate(class2 = as.factor(class2)) %>% 
#   mutate(species = fct_relevel(plot1_data$species))
# 
# ggplot(plot1_data, aes(species, r_fin)) +
#   geom_point(aes(color = class2))


# sort by variable and color by class

plot2_data = filled_data2 %>% 
  filter(temp_preferred != "NA") 

plot2_data = filled_data2

plot2_data <- plot2_data %>%  
  #mutate(length = sort(length)) %>% 
  mutate(species = as.factor(species)) %>% 
  mutate(class2 = as.factor(class2)) %>%
  mutate(order = as.factor(order)) %>% 
  mutate(species = fct_relevel(plot2_data$species))

#more descriptive**** # USE THESE PLOTS for continuous 
ind3 <- ggplot(plot2_data, aes(r_fin, class2)) +
  geom_boxplot(aes(color = class2)) +
  theme_bw(base_size  = 20) +
  labs(y = "Movement Class",
       x = "Growth Rate", color = "Class")
  #theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  #theme(legend.position = "none")
ind3


ggsave(ind3, file = paste0("ind3.png"), path = here::here("04_figs", "individual_plots"))


# #more to find groupings
# ggplot(plot1_data, aes(species, r_fin)) +
#   geom_point(aes(color = class2))
```


```{r}

plot2_data = filled_data2

plot2_data <- plot2_data %>%  
  #mutate(length = sort(length)) %>% 
  mutate(species = as.factor(species)) %>% 
  mutate(class2 = as.factor(class2)) %>%
  mutate(movement_keyword = as.factor(movement_keyword)) %>% 
  mutate(species = fct_relevel(plot2_data$species))

#more descriptive**** # USE THESE PLOTS for continuous 
ind2 <- ggplot(plot2_data, aes(movement_keyword, class2)) +
  geom_jitter(aes(color = class2)) +
  theme_bw(base_size  = 20) +
  labs(y = "Movement Class",
       x = "Movement Keyword", color = "Class") +
  theme(axis.text.x = element_text(angle = 30, hjust=1, size = 8))
  #theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  #theme(legend.position = "none")
ind2


ggsave(ind2, file = paste0("ind2.png"), path = here::here("04_figs", "individual_plots"))

```






```{r}
#categorical
plot2_data = filled_data %>% 
  select(demers_pelag, r_fin, species, class2, family, movement_keyword, 
         order, class_taxon, longevity_wild, diet_troph3) %>% 
  mutate(class2 = as.factor(class2), order = as.factor(order))

# #to find groupings
# ggplot(plot2_data, aes(class2, demers_pelag)) +
#   geom_point(aes(color = family)) +
#   theme(legend.position = "none")

#more descriptive*** - USE THIS FOR CATEGORICAL
ggplot(plot2_data, aes(class2, movement_keyword)) +
  #geom_point(aes(color = demers_pelag)) +
  geom_boxplot(aes(color = class2)) +
  theme_bw() +
  labs(x = "Movement Keyword",
       y = "Movement Class") +
  #theme(legend.position = "none") +
  scale_color_viridis_d() +
  theme(axis.text.x = element_text(angle = 30, hjust=1)) 
  #scale_color_manual(values = c("", ""))
```


### Metabolic data

```{r}

plot2_data = fishbase_data
plot2_data <- plot2_data[!is.na(plot2_data$oxygen_cons),]

ggplot(plot2_data, aes(oxygen_cons, class2)) +
  geom_point(aes(color = class2))

#option1 sort by oxygen_cons
plot2_data = plot2_data %>% 
  select(oxygen_cons, species, class2, genus, order, temp_preferred) %>% 
  mutate(oxygen_cons = sort(oxygen_cons)) %>% 
  mutate(species = as.factor(species)) %>% 
  mutate(class2 = as.factor(class2)) %>% 
  mutate(species = fct_relevel(plot2_data$species))

ggplot(plot2_data, aes(oxygen_cons, class2)) +
  geom_jitter(aes(color = oxygen_cons)) +
  theme_bw() +
  theme(legend.position="none") +
  theme(axis.text.x=element_blank()) +
  labs(x = "Species",
       y = "O2 Consumption Rate") # +
  #scale_color_viridis_d()

# no noteworthy groupings for class - try coloring by family/order


#option 2 sort by class
plot2_data = plot2_data %>% 
  select(oxygen_cons, species, class2, order, temp_preferred)%>% 
  mutate(class2 = sort(class2)) %>% 
  mutate(species = as.factor(species)) %>% 
  mutate(class2 = as.factor(class2)) %>% 
  mutate(species = fct_relevel(plot2_data$species))

#more descriptive****
ggplot(plot2_data, aes(temp_preferred, class2)) +
  geom_point(aes(color = class2)) +
  theme_bw() +
  labs(x = "Optimal Temp",
       y = "Movement Class") +
  theme(legend.position = "none")



```




```{r}
# 
# plot3_data = fishbase_data %>% 
#   select(demers_pelag, env_temp, body_shape_i, species, class, class2) %>% 
#   mutate(class = as.factor(class), class2 = as.factor(class2))
# 
# #to find groupings
# ggplot(plot3_data, aes(species, env_temp)) +
#   geom_point(aes(color = class2))
# 
# #more descriptive
# ggplot(plot3_data, aes(body_shape_i, class2)) +
#   geom_jitter(aes(color = body_shape_i)) +
#   theme_bw() +
#   labs(x = "env_temp",
#        y = "Movement Class",
#        color = "Habitat Type") +
#   scale_color_viridis_d() +
#   theme(axis.text.x = element_text(angle = 30, hjust=1)) 
#   #scale_color_manual(values = c("", ""))
# 
# plot4_data = fishbase_data %>% 
#   select(length, species, class2) %>% 
#   #mutate(class2 = sort(length)) %>% this line is causing problemssssss
#   mutate(species = as.factor(species)) %>% 
#   mutate(class2 = as.factor(class2)) %>% 
#   mutate(species = fct_relevel(plot4_data$species))
# 
# ggplot(plot4_data, aes(species, length)) +
#   geom_jitter(aes(color = class2))
# 

```

### DO NMDS HERE INSTEAD OF PCA WITH EUCLIDEAN DISTANCE FROM NOW ON 


Most of this can probably go in old now and then doing overall NMDS and FAMD then subset ones

### PCA with new variables added

```{r}
fishbase_p1 <- fishbase_data %>% 
  select(where(is.numeric)) %>% 
  select(r_fin, longevity_wild, length, fecundity_mean, diet_troph3, class2) %>% 
  na.omit() # YOU WANT TO REMOVE COLUMNS WITH NAS FOR PCA, AND PROBABLY NMDS TOO
 
p1 <- PCA(fishbase_p1[,1:5], graph = FALSE) # runs the PCA on the first 5 columns (not class)
p1

eigen_p1 <- p1$eig
eigen_p1 # 69.07% variance explained by first 2 components

p1_fig <- fviz_pca_ind(p1, label = "none", habillage = fishbase_p1$class2, addEllipses = TRUE, 
                ellipse.level = 0.95) +
            theme_minimal()
p1_fig

ggsave(p1_fig, file = paste0("p1_fig.png"), path = here::here("04_figs", "current_pca_famd"))


# Vizualize variable contributions to components

fviz_pca_biplot(p1, repel = TRUE, label = "none", habillage = fishbase_p1$class2)


# Contributions of variables to PC1
p1_c1 <- fviz_contrib(p1, choice = "var", axes = 1, top = 10)

# Contributions of variables to PC2
p1_c2 <- fviz_contrib(p1, choice = "var", axes = 2, top = 10)

components_p1 <- gridExtra::grid.arrange(p1_c1, p1_c2, ncol = 2)


#ggsave(components_p1, file = paste0("components_p1.png"), path = here::here("04_figs", "current_pca_famd"))

```



```{r}
fishbase_p2 <- fishbase_data %>% 
  select(where(is.numeric)) %>% 
  select(r_fin, length, diet_troph3, class2) %>% 
  mutate(class2 = as.factor(class2)) %>% 
  na.omit()

p2 <- PCA(fishbase_p2[,1:3], graph = FALSE) # runs the PCA on the first 4 columns (not class)
p2

eigen_p2 <- p2$eig
eigen_p2 # 74.71% of variance explained by first 2 components

p2_fig <- fviz_pca_ind(p2, label = "none", habillage = fishbase_p2$class2, addEllipses = TRUE, 
                ellipse.level = 0.95) +
            theme_minimal()
p2_fig

#ggsave(p2_fig, file = paste0("p2_fig.png"), path = here::here("04_figs"))

```



### FAMD with new variables added

```{r}

fishbase_f1 <- fishbase_data %>% 
   select(weight, env_temp, temp_preferred, fecundity_mean, r_fin, class2) %>% 
   mutate(class2 = as.factor(class2), 
          env_temp = as.factor(env_temp)) %>% 
   na.omit()
 
f1 <- FAMD(fishbase_f1[,1:5], graph=FALSE)
f1

eigen_f1 <- f1$eig
eigen_f1 # 42.11% of variance explained by first 2 components


f1_fig <- fviz_famd_ind(f1, label = "none", habillage = fishbase_f1$class2, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

f1_fig

# ggsave(f1_fig, file = paste0("f1_fig.png"), path = here::here("04_figs"))


```

```{r}

fishbase_f2 <- fishbase_data %>% 
   select(r_fin, env_temp, temp_preferred, demers_pelag, 
          body_shape_i, fecundity_mean, longevity_wild, movement_keyword, length, class2, kfin, diet_troph3, food_i) %>% 
   mutate(class2 = as.factor(class2), 
          env_temp = as.factor(env_temp),
          demers_pelag = as.factor(demers_pelag),
          body_shape_i = as.factor(body_shape_i)) %>% 
   na.omit()
 
f2 <- FAMD(fishbase_f2[,1:12], graph=FALSE)
f2

eigen_f2 <- f2$eig
eigen_f2 # 42.11% of variance explained by first 2 components


f2_fig <- fviz_famd_ind(f2, label = "none", habillage = fishbase_f2$class2, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

f2_fig

fviz_famd_var(f2, geom = "arrow", )


#ggsave(f2_fig, file = paste0("f1_fig.png"), path = here::here("04_figs"))


```


```{r}

fishbase_f3 <- fishbase_data %>% 
  select(length, diet_troph3, demers_pelag, order) %>% 
  mutate(order = as.factor(order), demers_pelag = as.factor(demers_pelag)) %>% 
  na.omit()

f3 <- FAMD(fishbase_f3[,1:3], graph=FALSE)
f3

eigen_f3 <- f3$eig
eigen_f3 # 30.38% of variance explained by first 2 components


f3_fig <- fviz_famd_ind(f3, label = "none", habillage = fishbase_f3$order, addEllipses = FALSE, ellipse.level = 0.95) +
  theme_minimal() +
  theme(legend.position = "none")
f3_fig 


#ggsave(f3_fig, file = paste0("f3_fig.png"), path = here::here("04_figs", "current_pca_famd"))

```






### Initial PCA and FAMD

```{r}
fishbase_pca1 <- fishbase_data %>% 
  select(where(is.numeric)) %>% 
  select(r_fin, kfin, length, diet_troph3, class) %>% 
  mutate(class = as.factor(class))

# this will only work if all variables are numeric
#cor.mat <- round(cor(fishbase_pca1),2) # nothing is highly correlated - good!
#cor.mat

pca1 <- PCA(fishbase_pca1[,1:4], graph = FALSE) # runs the PCA on the first 4 columns (not class)
pca1

eigen_pca1 <- pca1$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
eigen_pca1 # 66.4% of variance explained by first 2 components

fviz_screeplot(pca1, ncp=10) # scree plot shows eigenvalues/variances associated w components


plot(pca1, choix = "var") # this is the same plot as the original PCA code on line 31?


pca1_fig <- fviz_pca_ind(pca1, label = "none", habillage = fishbase_pca1$class, addEllipses = TRUE, 
                ellipse.level = 0.95) +
            theme_minimal()
pca1_fig

#ggsave(pca1_fig, file = paste0("pca1_fig.png"), path = here::here("04_figs"))

```


```{r}

fishbase_pca2 <- fishbase_data %>% 
  select(where(is.numeric)) %>% 
  select(r_fin, length, diet_troph3, class) %>% 
  mutate(class = as.factor(class))

pca2 <- PCA(fishbase_pca2[,1:3], graph = FALSE) # runs the PCA on the first 4 columns (not class)

eigen_pca2 <- pca2$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
eigen_pca2 # 81.45% of variance explained by first 2 components

plot(pca2, choix = "var")


pca2_fig <- fviz_pca_ind(pca2, label = "none", habillage = fishbase_pca2$class, addEllipses = TRUE, 
                ellipse.level = 0.95) +
            theme_minimal()
pca2_fig

#ggsave(pca2_fig, file = paste0("pca2_fig.png"), path = here::here("04_figs"))

```


```{r}

fishbase_pca3 <- fishbase_data %>% 
  select(where(is.numeric)) %>% 
  select(kfin, length, diet_troph3, class) %>% 
  mutate(class = as.factor(class))

pca3 <- PCA(fishbase_pca3[,1:3], graph = FALSE) # runs the PCA on the first 4 columns (not class)

eigen_pca3 <- pca3$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
eigen_pca3 # 81.29% of variance explained by first 2 components

plot(pca3, choix = "var")


pca3_fig <- fviz_pca_ind(pca3, label = "none", habillage = fishbase_pca3$class, addEllipses = TRUE, 
                ellipse.level = 0.95) +
            theme_minimal()
pca3_fig

#ggsave(pca3_fig, file = paste0("pca3_fig.png"), path = here::here("04_figs"))

```


```{r}

fishbase_pca4 <- fishbase_data %>% 
  select(where(is.numeric)) %>% 
  select(length, diet_troph3, class) %>% 
  mutate(class = as.factor(class))

pca4 <- PCA(fishbase_pca4[,1:2], graph = FALSE) # runs the PCA on the first 4 columns (not class)

eigen_pca4 <- pca4$eig # large eigenvalue (>1) for a component = it accounts for greater amt of variation than og components
eigen_pca4 # only 2 components

plot(pca4, choix = "var")


pca4_fig <- fviz_pca_ind(pca4, label = "none", habillage = fishbase_pca4$class, addEllipses = TRUE, 
                ellipse.level = 0.95) +
            theme_minimal()
pca4_fig # group 6 pretty clearly defined?

#ggsave(pca4_fig, file = paste0("pca4_fig.png"), path = here::here("04_figs"))

```


### FAMD

```{r}

fishbase_famd1 <- fishbase_data %>% 
  select(r_fin, kfin, length, diet_troph3, movement_keyword, food_i, class) %>% 
  mutate(class = as.factor(class), movement_keyword = as.factor(movement_keyword), food_i = as.factor(food_i)) %>% 
  na.omit(fishbase_famd1)


famd1 <- FAMD(fishbase_famd1[,1:6], graph=FALSE)
famd1

eigen_famd1 <- famd1$eig
eigen_famd1 # 7.23% of variance explained by first 2 components!?


famd1_fig <- fviz_famd_ind(famd1, label = "none", habillage = fishbase_famd1$class, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

famd1_fig
#ggsave(famd1_fig, file = paste0("famd1_fig.png"), path = here::here("04_figs"))


```


```{r}

fishbase_famd2 <- fishbase_data %>% 
  select(length, food_i, class) %>% 
  mutate(class = as.factor(class), food_i = as.factor(food_i)) %>% 
  na.omit(fishbase_famd2)

famd2 <- FAMD(fishbase_famd2[,1:2], graph=FALSE)

eigen_famd2 <- famd2$eig
eigen_famd2 # 5.49% 


famd2_fig <- fviz_famd_ind(famd2, label = "none", habillage = fishbase_famd2$class, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

famd2_fig
#ggsave(famd2_fig, file = paste0("famd2_fig.png"), path = here::here("04_figs"))


```


```{r}

fishbase_famd3 <- fishbase_data %>% 
  select(r_fin, kfin, diet_troph3, iucn_category, class) %>% 
  mutate(class = as.factor(class), iucn_category = as.factor(iucn_category)) %>% 
  na.omit(fishbase_famd3)

famd3 <- FAMD(fishbase_famd3[,1:4], graph=FALSE)
famd3

eigen_famd3 <- famd3$eig
eigen_famd3 # 31.50% of variance explained by first 2 components


famd3_fig <- fviz_famd_ind(famd3, label = "none", habillage = fishbase_famd3$class, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

famd3_fig 
#ggsave(famd3_fig, file = paste0("famd3_fig.png"), path = here::here("04_figs"))


```


```{r}

fishbase_famd4 <- fishbase_data %>% 
  select(length, iucn_category, class) %>% 
  mutate(class = as.factor(class), iucn_category = as.factor(iucn_category)) %>% 
  na.omit(fishbase_famd4)

famd4 <- FAMD(fishbase_famd4[,1:2], graph=FALSE)
famd4

eigen_famd4 <- famd4$eig
eigen_famd4 # 38.75% of variance explained by first 2 components


famd4_fig <- fviz_famd_ind(famd4, label = "none", habillage = fishbase_famd4$class, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()

famd4_fig
#ggsave(famd4_fig, file = paste0("famd4_fig.png"), path = here::here("04_figs"))


```


```{r}

fishbase_famd5 <- fishbase_data %>% 
  select(length, demers_pelag, diet_troph3, class) %>% 
  mutate(class = as.factor(class), 
         demers_pelag = as.factor(demers_pelag), 
         diet_troph3 = as.factor(diet_troph3)) %>% 
  na.omit(fishbase_famd5)

famd5 <- FAMD(fishbase_famd5[,1:3], graph=FALSE)
famd5

eigen_famd5 <- famd5$eig
eigen_famd5 # 1.27% of variance explained by first 2 components!?


famd5_fig <- fviz_famd_ind(famd5, label = "none", habillage = fishbase_famd5$class, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()
famd5_fig

ggsave(famd5_fig, file = paste0("famd5_fig.png"), path = here::here("04_figs"))

```


```{r}

fishbase_famd6 <- fishbase_data %>% 
  select(length, diet_troph3, demers_pelag, class) %>% 
  mutate(class = as.factor(class), demers_pelag = as.factor(demers_pelag)) %>% 
  na.omit()

famd6 <- FAMD(fishbase_famd6[,1:3], graph=FALSE)
famd6

eigen_famd6 <- famd6$eig
eigen_famd6 # 30.38% of variance explained by first 2 components


famd6_fig <- fviz_famd_ind(famd6, label = "none", habillage = fishbase_famd6$class, addEllipses = TRUE, ellipse.level = 0.95) +
  theme_minimal()
famd6_fig # still pretty good groupings here

famd6_1_fig <- fviz_famd_ind(famd6, label = "none", habillage = fishbase_famd6$class) +
  theme_minimal()
famd6_1_fig # really interesting.. what does it mean?


#ggsave(famd6_fig, file = paste0("famd6_fig.png"), path = here::here("04_figs"))
#ggsave(famd6_1_fig, file = paste0("famd6_1_fig.png"), path = here::here("04_figs"))

```






### Old


```{r}
# 
# fishbase_famd5 <- fishbase_data %>% 
#   select(r_fin, kfin, length, diet_troph3, movement_keyword, demers_pelag, food_i, class) %>% 
#   mutate(class = as.factor(class), 
#          movement_keyword = as.factor(movement_keyword), 
#          food_i = as.factor(food_i), 
#          demers_pelag = as.factor(demers_pelag)) %>%  # error here?  <--------------------------------------------
#   na.omit(fishbase_famd5)
#   
# 
# famd5 <- FAMD(fishbase_famd5[,1:7], graph=FALSE)
# famd5
# 
# eigen_famd5 <- famd5$eig
# eigen_famd5 # 7.23% of variance explained by first 2 components!?
# 
# 
# famd5_fig <- fviz_famd_ind(famd5, label = "none", habillage = fishbase_famd5$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
# 
# famd5_fig
# #ggsave(famd5_fig, file = paste0("famd5_fig.png"), path = here::here("04_figs"))
# 

```


```{r}
# 
# fishbase_famd5 <- fishbase_data %>% 
#   select(r_fin, kfin, length, diet_troph3, food_i, movement_keyword, iucn_category, demers_pelag, class) %>% 
#   mutate(class = as.factor(class), 
#          demers_pelag = as.factor(demers_pelag), 
#          iucn_category = as.factor(iucn_category), 
#          food_i = as.factor(food_i), 
#          movement_keyword = as.factor(movement_keyword)) %>% 
#   na.omit(fishbase_famd5) # error why?  <------------------------------------------------------------------
# 
# famd5 <- FAMD(fishbase_famd5[,1:8], graph=FALSE)
# famd5
# 
# eigen_famd5 <- famd5$eig
# eigen_famd5 # 30.38% of variance explained by first 2 components
# 
# 
# famd5_fig <- fviz_famd_ind(famd5, label = "none", habillage = fishbase_famd5$class, addEllipses = TRUE, ellipse.level = 0.95) +
#   theme_minimal()
# famd5_fig 
# 
# 
# #ggsave(famd5_fig, file = paste0("famd5_fig.png"), path = here::here("04_figs"))
# 

```
